PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-05 paper/sec-results.tex
Top pair, sect 3

\paragraph{Top-quark pair production.} We now consider the top-pair differential
distributions measured by the ATLAS and CMS experiments at
\SI{7}{\tera\electronvolt}. In figure~\ref{fig:atlastop} we report the
distributions differential in the transverse momentum of the top quark,
$p_\mathrm{T}^\mathrm{t}$, and in the invariant mass of the top-quark pair,
$m_{\mathrm{t}\bar{\mathrm{t}}}$. Analogous plots for the distributions differential
in the rapidity of either the top quark or the top-quark pair are collected in
appendix~\ref{app:add_plots}. From figure~\ref{fig:atlastop}, we immediately
validate the \textsc{PineAPPL} result: its relative difference with respect to
the MC result is at most as large as $0.4\tcperthousand$, irrespective of the
accuracy of the theory, of the choice of scale and of the distribution
considered.

@@

The process is dominated by

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMS_Z_13_TEV}
    \caption{PineAPPL comparison for CMS Z $p_\mathrm{T}$ distribution.}
    \label{fig:cmsZ13TeV}
\end{figure}

>>>>

\paragraph{Top-quark pair production.} We now consider the top-pair differential
distributions measured by the ATLAS and CMS experiments at
\SI{8}{\tera\electronvolt}. In figure~\ref{fig:atlastop} we report the
distributions differential in the transverse momentum of the top quark,
$p_\mathrm{T}^\mathrm{t}$, and in the invariant mass of the top-quark pair,
$m_{\mathrm{t}\bar{\mathrm{t}}}$. Analogous plots for the distributions differential
in the rapidity of either the top quark or the top-quark pair are collected in
appendix~\ref{app:add_plots}. From figure~\ref{fig:atlastop}, we immediately
validate the \textsc{PineAPPL} result: its relative difference with respect to
the MC result is at most as large as $0.4\tcperthousand$, irrespective of the
accuracy of the theory, of the choice of scale and of the distribution
considered.

@@

The process receives its leading contribution from the $\mathrm{gg}$ channel,
which varies between 81\% and 61\% (76\% and 83\%) of the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) differential cross section as the value of the
transverse momentum of the top quark (the invariant mass of the top-quark pair)
increases in the bin range; $\mathrm{\gamma}\mathrm{g}$ scattering
correspondingly accounts for about 0.5\% -- 1\% (0.5\% -- 0.7\%) of the cross
section; the contribution from other PI parton luminosities is comparatively
negligible. Overall, the EW corrections suppress the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) distribution by about 0.2\% -- 3.5\%
(0.5\% -- 0.2\%) for increasing values of $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$), except in the first bin of the
$p_\mathrm{T}^\mathrm{t}$ distribution, where they enhance the cross section by
about 1\%. The size of these shifts, however, remains always significantly
smaller than the data uncertainty.\footnote{In
  figure~\ref{fig:atlastop} the data uncertainty corresponds to the ATLAS
  measurement~\cite{Aad:2015auj}. Similar considerations apply also for the CMS
  measurement~\cite{Khachatryan:2015oaa}.}
As a consequence, we anticipate that the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty, should the data
be fitted with the inclusion of EW corrections; PDFs will possibly not become
more precise.

The size of the EW correction is comparable to the size of the PDF uncertainty,
except at large values of transverse momentum or invariant mass, where the
former becomes larger than the latter. This fact suggests that, once included in
a global fit, EW corrections can improve the accuracy of the PDFs. In comparison
to the scale uncertainty, the size of the EW corrections remains negligible:
despite the fact that the choice of factorisation and renormalisation scales
have been devised to optimise the convergence of the perturbative expansion,
NNLO QCD corrections remain large, as expected in a process mostly initiated
by gluons. Their inclusion is therefore mandatory in a fit of PDFs. Finally,
the Monte Carlo statistical uncertainty remains negligible in comparison to the
data, PDF and scale uncertainties, and to the size of the EW correction. Our
conclusions are therefore not affected by Monte Carlo inefficiencies.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMS_Z_13_TEV}
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb} but for the 
      CMS differential Z $p_\mathrm{T}$ measurement at a centre-of-mass energy of
      \SI{13}{\tera\electronvolt}~\cite{Sirunyan:2019bzr}.}
    \label{fig:cmsZ13TeV}
\end{figure}
%-------------------------------------------------------------------------------
